---
title: "Bus Commute Reliability Report"
output: html_fragment
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, error=FALSE)

# Load necessary library
library(tidyverse)

# Import and clean data
csv_url <- "https://docs.google.com/spreadsheets/d/e/2PACX-1vQnq8A60TB2xCDycdfxkTKL3JHKT9IARTOwnv97DtN3_tIYzqKZd3hvrZzaP2ajNVP7OnkBVO7kGE5e/pub?gid=1834374921&single=true&output=csv"

logged_data <- read_csv(csv_url)

latest_data <- logged_data %>%
  rename(bus_capacity = 2,
         delay_mins = 3,
         bus_route = 4)

# Calculate summary values for dynamic comments
average_delay <- mean(latest_data$delay_mins, na.rm = TRUE)
max_delay <- max(latest_data$delay_mins, na.rm = TRUE)
total_logs <- nrow(latest_data)
```

```{css}
/* Optional styling for your dynamic comments */
.dynamic-comment {
  background-color: #f0f8ff;
  padding: 15px;
  border-left: 5px solid #4682b4;
  font-style: italic;
}
```

## Analysis of Daily Bus Commute Reliability

This report provides a dynamic overview of public transport performance based on observational logging. Over the past two weeks, I have recorded data on bus arrival delays and passenger capacity across various routes to identify patterns in service reliability.

::: dynamic-comment
**Key Insights:**  
To date, a total of **`r total_logs`** observations have been recorded. On average, buses arrived **`r round(average_delay, 1)`** minutes after their scheduled time. The most significant delay observed during this period was **`r max_delay`** minutes.
:::

### Observed Bus Capacity
The chart below visualises how crowded the buses were during the observed trips. This categorical data helps in understanding peak-time pressure on specific routes.

```{r fig.width=8, fig.height=4}
# Chart 1: Distribution of observed Bus Capacity
latest_data %>%
  ggplot() +
  geom_bar(aes(x = bus_capacity), fill = "steelblue") +
  theme_minimal() +
  labs(title = "Frequency of Bus Capacity Levels",
       subtitle = "Categorical distribution of passenger volume",
       x = "Capacity Description",
       y = "Number of Observations")
```

### Route Distribution
This visualisation shows the frequency of logs per bus route, providing context on which services were monitored most frequently during the data collection phase.

```{r fig.width=8, fig.height=4}
# Chart 2: Frequency of specific Bus Routes logged
latest_data %>%
  ggplot() +
  geom_bar(aes(x = bus_route), fill = "darkorange") +
  theme_minimal() +
  labs(title = "Summary of Bus Routes Logged",
       subtitle = "Distribution of transport routes tracked",
       x = "Bus Route Number",
       y = "Frequency")
```